Broom: Safe Region-Based Memory Management for Dataflow Programs

### 2015-11-03 paper/type-inference.tex
Edits.

Region type templates are FGJ types extended with fresh region
variables ($\rho$) and predicate variables ($\varphi$) to denote
unknown region annotations and region constraints, respectively. For
instance, if a variable $x$ has type $\ObjZ$ in FGJ, its region type
template is of form $\ObjZ\inang{\rho_0}$, where $\rho_0$ is a fresh
region variable. Likewise, given the region-annotated definition of
\C{Pair} class from \S\ref{sec:fb-syntax} a region type template for
the method \C{mkSnd} shown below is \footnote{In our
exposition, we assume that classes $\C{A}$ and $\C{B}$ are trivial
subclasses of $\ObjZ$ with no fields/methods. Like $\ObjZ$, they
accept one region parameter - the allocation region of their
objects.}\footnote{We abuse arrow notation to also represent types of
methods, but unlike function types, there is no allocation region
annotation atop the arrow in a method type.}
$\inang{\rhoalloc_0,\rho_1,\rho_2,\rho_3 \,|\, \varphi_0}
\C{Pair}\inang{\C{A},\C{B}} \inang{\rho_1,\rho_2,\rho_3} \rightarrow
\unitZ$, where $\rhoalloc_0$ and $\rho_{1-3}$ are fresh region
variables, and $\varphi_0$ is a fresh predicate variable denoting
unknown constraints over $\rhoalloc_0$ and $\rho_{1-3}$.  
\begin{codejava}
  unit mkSnd(Pair<A,B> p) {
    return (p.snd = new B());
  }
\end{codejava}
Region type template of a type variable $\C{a}$ is $\C{a}@\rho$, where
$\rho$ is fresh. For a class, the template is derived by generalizing
free region variables in the templates of its fields as region
parameters, and introducing a new predicate variable to denote
unknown constraints on region parameters. A fresh region variable
denoting allocation region argument is also added as a parameter. For
example, given the standard FGJ definition of 
$\C{Pair}\inang{\C{a},\C{b}}$ class, if $\C{a}$
and $\C{b}$, the types of $\C{fst}$ and $\C{snd}$ fields (resp.), are
templatized as $\C{a}@\rho_1$ and $\C{b}@\rho_2$ (resp.), then
the template for the class in \FB is:
\begin{center}
$\C{Pair}\inang{\C{a}\extends\ObjZ,\C{b}\extends\ObjZ}
\inang{\rhoalloc_0,\rho_0,\rho_1\,|\,\varphi_0} \extends \ObjZ \{...\}$
\end{center}
Fresh region variable $\rhoalloc_0$ is added to denote the allocation
region argument. The body of the class, which is elided with the
ellipsis above, contains region type templates of all its fields,
methods and constructors.

@@

Function $\elabExpr$, shown in Fig.~\ref{fig:fb-elabexpr}, performs
this elaboration for a subset of expressions in $\absof{\FB}$, whose
corresponding $\FB$ expressions have been ascribed static semantics in
Fig.~\ref{fig:fb-staticsem}. $\elabExpr$ is defined under the same
context as the expression typing judgment in
Fig.~\ref{fig:fb-staticsem} with symbols $\A$,$\ralloc$, and $\env$
retaining their meaning. The function traverses expressions in a
syntax-directed manner of a type checker, introducing fresh region
type templates for unknown region types, while generating constraints
over region and predicate variables. The precise nature of these
constraints will be clarified shortly, but their essential nature is
that they capture the relationship between the type templates of
various subexpressions and their well-formedness. Note that
$\elabExpr$ returns the region type template of the subexpression,
which is used to generate constraints for the expression. Functions
$\typeOk$ and $\subtypeOk$ (definitions not shown) used by $\elabExpr$
implement type well-formedness and subtype judgments from
Fig.~\ref{fig:fb-staticsem}, respectively.

@@

$\elabClass$ elaborates a class's constructor ($k$) and method
definitions ($\bar{d}$) using $\elabCons$\footnote{The definition of
$\elabCons$ is straightforward, hence not shown.} and $\elabMeth$
functions, respectively. To the set of constraints returned by these
functions, $\elabClass$ adds constraints generated by checking the
well-formedness of the type templates of its superclass and fields,
and also a new constraint capturing the safety condition that
allocation regions of objects referred by the instance variables
should outlive the allocation region of the instance itself.

@@

@@

Constraints generated by the method elaboration algorithm
($\elabMeth$) are of two kinds: 
\begin{itemize}
\item Well-formedness constraints over region variables ($\rho$) of
form $\rho \in \rhoenv$, where $\rhoenv$ is a set of region
identifiers, and 
\item Validity constraints of form $\isvalid{\phicx}{\phi}$, where
$\phicx$ and $\phi$ are conjunctions of (a). outlives and equality
constraints over region variables and region identifiers, and (b).
a predicate variable ($\varphi$) denoting the unknown pre-condition
of the method being elaborated.
\end{itemize}
\noindent The task of a constraint solving algorithm is to assign
regions to region

>>>>

Region type templates are FGJ types extended with fresh region
variables ($\rho$) and predicate variables ($\varphi$) to denote
unknown region annotations and region constraints, respectively. For
instance, if a variable $x$ has type $\ObjZ$ in FGJ, its region type
template is of form $\ObjZ\inang{\rho_0}$, where $\rho_0$ is a fresh
region variable. Likewise, given the region-annotated definition of
\C{Pair} class from \S\ref{sec:fb-syntax} a region type template for
the method \C{mkSnd} shown below is \footnote{In our
exposition, we assume that classes $\C{A}$ and $\C{B}$ are trivial
subclasses of $\ObjZ$ with no fields/methods. Like $\ObjZ$, they
accept one region parameter - the allocation region of their
objects.}\footnote{We abuse arrow notation to also represent types of
methods, but unlike function types, there is no allocation region
annotation atop the arrow in a method type.}
$\inang{\rhoalloc_0,\rho_1,\rho_2,\rho_3 \,|\, \varphi_0}
\C{Pair}\inang{\C{A},\C{B}} \inang{\rho_1,\rho_2,\rho_3} \rightarrow
\unitZ$, where $\rhoalloc_0$ and $\rho_{1-3}$ are fresh region
variables, and $\varphi_0$ is a fresh predicate variable denoting
unknown constraints over $\rhoalloc_0$ and $\rho_{1-3}$.  
\begin{codejava}
  unit mkSnd(Pair<A,B> p) {
    return (p.snd = new B());
  }
\end{codejava}
Region type template of a type variable $\C{a}$ is $\C{a}@\rho$, where
$\rho$ is fresh. For a class, a template is computed in two steps. In
the first step, we templatize the types of all its fields, constructor
arguments, and arguments and return values of all its methods, along
with the type of its superclass. In the second step, we generalize all
the free region variables occuring in the templates computed in the
first step as region parameters of the class. Finally, we add a fresh
allocation region parameter ($\rhoalloc$) to the list of parameters,
and introduce a new predicate variable ($\varphi$) to denote unknown
constraints on region parameters. For example, consider the standard
FGJ definition of $\C{Pair}\inang{\C{a},\C{b}}$ class, where $\C{a}$
and $\C{b}$ are the types of $\C{fst}$ and $\C{snd}$ fields,
respectively. It can be templatized as following:
\begin{center}
\begin{codejava}
class Pair$\inang{a\extends\ObjZ, b\extends\ObjZ}$
          $\inang{\rhoalloc_0,\rho_{0-4}\,|\,\varphi_0} \extends \ObjZ\inang{\rho_4}$ {
  $a@\rho_0$ fst;
  $b@\rho_1$ snd;
  Pair($a@\rho_2$ fst, $b@\rho_3$ snd) { $...$ }
  $...$
}
\end{codejava}
\end{center}
% \footnote{In reality, the generalized region variables only come
% from the type templates of class's fields and superclass. Method
% types do not contain any free region variables.}
The template elides $\C{Pair}$'s methods, whose region type templates
contain no free variables. Among the region parameters of the class
template, $\rho_{0-4}$ are obtained by generalizing free region
variables in the types of its class's fields, constructor arguments,
and its superclass type.  The remaining parameter ($\rhoalloc_0$) is a
fresh region variable denoting the allocation region argument. Fresh
predicate variable $\varphi_0$ denotes unknown constraints over
$\rhoalloc_0$ and $\rho_{0-4}$ that need to hold for template to be a
well-formed region-annotated class definition in $\FB$.

@@

Function $\elabExpr$, shown in Fig.~\ref{fig:fb-elabexpr}, performs
this elaboration for a subset of expressions in $\absof{\FB}$, whose
corresponding $\FB$ expressions have been ascribed static semantics in
Fig.~\ref{fig:fb-staticsem}. $\elabExpr$ is defined under the same
context as the expression typing judgment in
Fig.~\ref{fig:fb-staticsem} with symbols $\A$,$\ralloc$, and $\env$
retaining their meaning. The function traverses expressions in a
syntax-directed manner of a type checker, introducing fresh region
type templates for unknown region types, while generating constraints
over region and predicate variables. The precise nature of generated
constraints is explained in \S\ref{sec:fb-constraintsem}, but in
summary, they capture the relationships between the type templates of
various subexpressions and their well-formedness. Note that
$\elabExpr$ returns the region type template of the subexpression,
which is used to generate constraints for the expression. Functions
$\typeOk$ and $\subtypeOk$ (definitions not shown) used by $\elabExpr$
implement type well-formedness and subtype judgments from
Fig.~\ref{fig:fb-staticsem}, respectively.

@@

$\elabClass$ elaborates the definition of a class $B$. It relies on
$\elabCons$\footnote{The definition of $\elabCons$ is straightforward,
hence not shown.} and $\elabMeth$ functions to elaborate $B$'s
constructor ($k$) and method definitions ($\bar{d}$), respectively. To
the set of constraints returned by these functions, $\elabClass$ adds
constraints generated by checking the well-formedness of the type
templates of its superclass and fields, and also a new constraint
capturing a couple of safety conditions: first, the allocation regions
of objects referred by the instance variables should outlive the
allocation region of the instance itself, and second, the allocation
regions of a class type and its superclass type must be the same.

@@

\subsection{The Constraint Satisfaction Problem}
\label{sec:fb-constraintsem}

Our constraint generation algorithm generates three kinds of
constraints:
\begin{itemize}
\item Well-formedness constraints of form $\rho \in \rhoenv$,
restricting the domain of unification for a region variable ($\rho$)
to the set ($\rhoenv$) of regions in scope,
\item Well-formedness constraints of form
$\tywf{\rhoenv}{\varphi}$, restricting the domain of a predicate
variable ($\varphi$) to the set of all possible constraint formulas
over region variables ($\rhoenv$) in scope, and
\item Validity constraints of form $\isvalid{\phicx}{\phi}$, where
$\phicx$ and $\phi$ are region constraints (Fig.~\ref{fig:fb-syntax})
extended with predicate variables and \emph{pending
substitutions}\footnote{we borrowed this terminology
from~\cite{ltpldi08}}:
\begin{smathpar}
\begin{array}{lcl}
\phicx,\phi & \coloneqq & true \ALT \rho \outlives \rho \ALT \rho = \rho 
    \ALT F(\varphi) \ALT \phi \conj \phi\\
F & \coloneqq & \cdot \ALT [\rho/\rho]F \\
\end{array}
\end{smathpar}
\end{itemize}
A pending substitution ($F$) is a substitution function over region
variables/identifiers. They represent the substitutions that need to
be carried out when a predicate variable ($\varphi$) is replaced by a
concrete formula in a validity constraint. For instance, in the
validity constraint $\isvalid{\rgn_1 \outlives
\rgn_2}{[\rgn_1/\rho_1][\rgn_2/\rho_2]\varphi}$, pending substitution
is $[\rgn_1/\rho_1][\rgn_2/\rho_2]$. Any concrete formula (call it
$\phisol$) over variables $\rho_1$ and $\rho_2$ is a solution to
$\varphi$ if and only the formula obtained by substituting $\rgn_1$
and $\rgn_2$ for $\rho_1$ and $\rho_2$ (resp.) in $\phisol$ is
deducible from $\rgn_1 \outlives \rgn_2$.

% Validity constraints primarily result from checking if a region type
% template of form $B\inang{\tbar}\inang{\rhoalloc,\rhobar}$ is
% well-formed (i.e., $\typeOk$). The corresponding well-formedness rule
% (Fig.~\ref{fig:fb-staticsem}) requires that the instantiation of $B$'s
% formal region parameters with actual region arguments satisfy its
% precondition, where actuals are subsituted for formals. Since $B$'s
% precondition could be unknown, the requirement results in a in a
% validity constraint with predicate variables and pending
% substitutions. Other sources of validity constraints are calls to
% region-polymorphic methods, where similar situation arises. 
In general, validity constraints generated by our algorithm are of the
form:
\begin{center}
\( \isvalid{\phictxt \conj \bigwedge_{i}\varphi_i}
           {\phicstr \conj \bigwedge_{j}F_j(\varphi_j)}\)
\end{center}
Where $\varphi_i$'s denote the unknown preconditions of the class or
the method under elaboration. If the constraint is generated while
checking the well-formedness of a type or elaborating an expression,
then $\varphi_j$'s denote the unknown preconditions of classes and
methods that were used in that type or expression.  Each use of a
(region-polymorphic) class or a method may instantiate region
parameters differently, resulting in a different pending substitution
($F_j$). Formulas $\phictxt$ and $\phicstr$ are concrete, i.e., free
of predicate variables and pending substitutions.  They capture the
concrete relationships between regions that are \emph{known} to hold,
and \emph{required} to hold, respectively, under the context in which
the constraint was generated.

\paragraph{Constraints Example 1} The original source of all
outlives constraints is the validity constraint $C_3$ in $\elabClass$
function (Fig.~\ref{fig:fb-elabmeth}). In context of the $\C{Pair}$
class template from \S\ref{sec:fb-templatization}, the constraint is
as following:
\begin{center}
  \(\isvalid{\varphi_0}{\rho_0 \succeq \rhoalloc_0 \conj \rho_0
  \succeq \rhoalloc_0 \conj \rho_4 = \rhoalloc_0}\)
\end{center}

\paragraph{Constraints Example 2} Consider a contrived method
$\C{alt}$ that accepts an argument $\C{p}$ of type
$\C{Pair}\inang{\C{A},\C{A}}$, and two objects ($\C{x}$ and $\C{y}$)
of type $A$. It then assigns one object to $\C{p.fst}$ and other to
$\C{p.snd}$, and calls itself recursively without ever terminating.
Which one is assigned to $\C{fst}$, and which to $\C{snd}$, is
alternated between recursive calls. The region-annotated definition of
$\C{alt}$ is shown below:
\begin{codejava}
unit alt<$\rhoalloc$,$\rho_{0-4}$ | $\varphi_0$>(Pair<A,A><$\rho_0$,$\rho_1$,$\rho_2$> p,
                          A<$\rho_3$> x, A<$\rho_4$> y) {
  return p.fst = x; p.snd = y; 
         this.alt<$\rhoalloc$,$\rho_1$,$\rho_2$,$\rho_4$,$\rho_3$>(A,y,x)
}
\end{codejava}
Note that, to avoid clutter, we have already resolved appropriate
region arguments to the recursive call, instead of introducing new
region variables and generating equality constraints on them. Rest of
the validity constraints generated while elaborating $\C{alt}$ to the
above region-annotated definition are shown below:
\begin{center}
\(\isvalid{\varphi_0}{\rho_1 \outlives \rho_0 \conj \rho_2 \outlives \rho_0}\)
$\quad$
\(\isvalid{\varphi_0}{\rho_1 = \rho_3}\)
$\quad$
\(\isvalid{\varphi_0}{\rho_2 = \rho_4}\)\\
\(\isvalid{\varphi_0}{[\rho_3/\rho_4][\rho_4/\rho_3]\varphi_0}\)
\end{center}
The first constraint is generated by $\typeOk$ on the type of $\C{p}$.
Second and third are generated by the the assignment expressions, and
the last constraint is generated from the recursive call.

@@

### 2015-11-03 paper/type-system.tex
Edits.

Class definitions in \FB extend FGJ's with region parameters
($\rho$). Informally, region variables ($\rho$) are to region
identifiers ($\rgn$) as type variables ($a$) are to core types ($T$);
they can be instantiated with region identifiers, as needed. A class
in \FB is necessarily parameterized over the allocation region of its
objects. We use $\rho^a$ to denote this parameter, and often refer to
it as \emph{class}'s allocation region, although it is really the
allocation region of its objects. An Object in \FB can contain fields
referring to objects allocated in regions ($\rhobar$) other than its
own allocation region ($\rhoalloc$), provided that the former outlive
the later (i.e., $\rhobar \outlives \rho$). In such cases, the
definition of object's class needs to be parametric over allocation
regions of its fields (i.e., their classes). Furthermore, the
constraint that such regions must outlive the allocation region of the
class needs to be made explicit in the definition. For instance, a
\C{Pair} object allocated in a region ${\rhoalloc}$ can have its first
component coming from ${\rho_1}$ and second from $\rho_2$, where all
the $\rho$'s are distinct, and $\rho_1, \rho_2$ outlive
$\rhoalloc$. The corresponding definition of the \C{Pair} class is as
follows(The symbol $\extends$ should be read \emph{extends}):  
\begin{codejava}[mathescape=true]
class Pair<a $\extends$ Object, b $\extends$ Object>
          <$\rho^a, \rho_1, \rho_2 \,|\, \rho_1 \outlives \rho^a \conj \rho_2 \outlives \rho^a$>
           {
  a@$\rho_1$ fst; 
  b@$\rho_2$ snd;
  Pair(a@$\rho_1$ fst, b@$\rho_2$ snd) {
    super(); 
    this.fst = fst; 
    this.snd = snd;
  }
  a@$\rho_1$ getFst() {
    return this.fst;
  }
}
\end{codejava}
To construct objects of the \C{Pair} class, its type and region
parameters need to be instantiated with core types ($T$) and concrete
region identifiers ($\rgn$), respectively. For example:
\begin{codejava}
letregion $\rgn_0$ {
  Object<$\rgn_0$> snd = new Object<$\rgn_0$>();
  letregion $\rgn_1$ {
    Object<$\rgn_1$> fst = new Object<$\rgn_1$>();
    let p = new Pair<Object,Object><$\rgn_1$,$\rgn_0$,$\rgn_1$>
                  (fst,snd);
  }
}
\end{codejava}
In the above code, the instantiation of $\rhoalloc$ and $\rho_1$ with
$\rgn_0$, and $\rho_2$ with $\rgn_1$ is allowed because (a) $\rgn_0$
and $\rgn_1$ are live during the instantation, and (b). $\rgn_0
\outlives \rgn_1$ and $\rgn_1 \outlives \rgn_1$ (since outlives is
reflexive). Observe that the region type of \C{p} conveys
the fact that (a). it is allocated in region $\rgn_1$, and (b). it
holds references to objects allocated in region $\rgn_0$ and $\rgn_1$.
In contrast, if we choose to allocate the \C{snd} object also in
$\rgn_1$, the region type of \C{p} would be
\C{Pair<\ObjZ,\ObjZ><$\rgn_1$,$\rgn_1$,$\rgn_1$>}, which we abbreviate
as \C{Pair<\ObjZ,\ObjZ>@$\rgn_1$}\footnote{In general, we treat
$B\inang{\tbar}@\rgn$ as an abbreviation of
$B\inang{\tbar}\inang{\rbar}$}. The $\ObjZ$ class in \FB has no
fields, hence its region type is of form $\ObjZ@\rgn$.
% In general, the \C{@} notation in a
% region type of an object \C{x} highlights that
% \C{x}, and all the objects reachable from \C{x} via references are
% allocated in a single region. We say that \C{x} is \emph{contained} in
% the region. The $\ObjZ$ class in \FB contains no references to other
% objects, hence its objects are always contained in their allocation
% region. Their region type is $\ObjZ\inang{\rgn}$ (or equivalently,
% $\ObjZ@\rgn$), for some region $\rgn$.

@@

\begin{codejava}
class Region<a $\extends$ Object><${\rhoalloc}$> {
  <${\rho}$>unit$\xrightarrow{\rhoalloc}$a@$\rho$ thunk;
  Region(<${\rho}$>unit$\xrightarrow{\rhoalloc}$a@$\rho$ thunk) {
    super(); 
    this.thunk = thunk;
  }
  unit transfer$\inang{\rho}$(unit u) {
    return u;
  }
}
\end{codejava}

>>>>

Class definitions in \FB extend FGJ's with region parameters
($\rho$). Informally, region variables ($\rho$) are to region
identifiers ($\rgn$) as type variables ($a$) are to core types ($T$);
they can be instantiated with region identifiers, as needed. A class
in \FB is necessarily parameterized over the allocation region of its
objects. We use $\rho^a$ to denote this parameter, and often refer to
it as \emph{class}'s allocation region, although it is really the
allocation region of its objects. An Object in \FB can contain fields
referring to objects allocated in regions ($\rhobar$) other than its
own allocation region ($\rhoalloc$), provided that the former outlive
the later (i.e., $\rhobar \outlives \rho$). In such cases, the
definition of object's class needs to be parametric over allocation
regions of its fields (i.e., their classes). Furthermore, the
constraint that such regions must outlive the allocation region of the
class needs to be made explicit in the definition. For instance, a
\C{Pair} object allocated in a region ${\rhoalloc}$ can have its first
component coming from ${\rho_1}$ and second from $\rho_2$, where all
the $\rho$'s are distinct, and $\rho_1, \rho_2$ outlive
$\rhoalloc$. The corresponding definition of the \C{Pair} class is as
follows(The symbol $\extends$ should be read \emph{extends}):  
\begin{codejava}[mathescape=true]
class Pair<a $\extends$ Object, b $\extends$ Object>
          <$\rho^a, \rho_1, \rho_2 \,|\, \rho_1 \outlives \rho^a \conj \rho_2 \outlives \rho^a$> $\extends$ Object<$\rho^a$>
           {
  a@$\rho_1$ fst; 
  b@$\rho_2$ snd;
  Pair(a@$\rho_1$ fst, b@$\rho_2$ snd) {
    super(); 
    this.fst = fst; 
    this.snd = snd;
  }
  a@$\rho_1$ getFst() {
    return this.fst;
  }
}
\end{codejava}
To construct objects of the \C{Pair} class, its type and region
parameters need to be instantiated with core types ($T$) and concrete
region identifiers ($\rgn$), respectively. For example:
\begin{codejava}
letregion $\rgn_0$ {
  Object<$\rgn_0$> snd = new Object<$\rgn_0$>();
  letregion $\rgn_1$ {
    Object<$\rgn_1$> fst = new Object<$\rgn_1$>();
    let p = new Pair<Object,Object><$\rgn_1$,$\rgn_0$,$\rgn_1$>
                  (fst,snd);
  }
}
\end{codejava}
In the above code, the instantiation of $\rhoalloc$ and $\rho_1$ with
$\rgn_0$, and $\rho_2$ with $\rgn_1$ is allowed because (a) $\rgn_0$
and $\rgn_1$ are live during the instantation, and (b). $\rgn_0
\outlives \rgn_1$ and $\rgn_1 \outlives \rgn_1$ (since outlives is
reflexive). Observe that the region type of \C{p} conveys
the fact that (a). it is allocated in region $\rgn_1$, and (b). it
holds references to objects allocated in region $\rgn_0$ and $\rgn_1$.
In contrast, if we choose to allocate the \C{snd} object also in
$\rgn_1$, the region type of \C{p} would be
\C{Pair<\ObjZ,\ObjZ><$\rgn_1$,$\rgn_1$,$\rgn_1$>}, which we abbreviate
as \C{Pair<\ObjZ,\ObjZ>@$\rgn_1$}\footnote{In general, we treat
$B\inang{\tbar}@\rgn$ as an abbreviation of
$B\inang{\tbar}\inang{\rbar}$}. The $\ObjZ$ class in \FB has no
fields, hence its region type is of form $\ObjZ@\rgn$.
% In general, the \C{@} notation in a
% region type of an object \C{x} highlights that
% \C{x}, and all the objects reachable from \C{x} via references are
% allocated in a single region. We say that \C{x} is \emph{contained} in
% the region. The $\ObjZ$ class in \FB contains no references to other
% objects, hence its objects are always contained in their allocation
% region. Their region type is $\ObjZ\inang{\rgn}$ (or equivalently,
% $\ObjZ@\rgn$), for some region $\rgn$.

@@

\begin{codejava}
class Region<a $\extends$ Object><${\rhoalloc}$> $\extends$ Object<$\rhoalloc$>{
  <${\rho}$>unit$\xrightarrow{\rhoalloc}$a@$\rho$ thunk;
  Region(<${\rho}$>unit$\xrightarrow{\rhoalloc}$a@$\rho$ thunk) {
    super(); 
    this.thunk = thunk;
  }
  unit transfer$\inang{\rho}$(unit u) {
    return u;
  }
}
\end{codejava}
